Adaptive Safe Strong Screening for Efficient Lasso-Type Problems Optimization over the Solution Path

### 2021-01-12 main.tex
Edits to 3.1

\subsection{Safe rules}

@@

\pb{Mention Sure Iterative Screening and why we do not focus on it here; maybe in general, just introduce the idea of stochastic screening}

\section{Adaptive hybrid rule}
\label{sec:method}

In this section, we will introduce the framework of adaptive hybrid rules, implement them with EDPP as the safe rule to form Ada-SSR-EDPP and propose a strategy to adaptively updating the reference.

\subsection{Adaptive safe rule}

From the SEDPP rule \eqref{eq:sedpp} in Theorem 2.2, we can do screening at $\lambda_{k+1}$ using the solution at $\lambda_k$ as a reference. Now consider we are solving the lasso problem along the path $\lambda_0,...,\lambda_k,\lambda_{k+2},...,\lambda_K$ instead. Using the SEDPP rule, we can do screening at $\lambda_{k+2}$ using the solution at $\lambda_k$ as a reference as well, which is considered the previous solution now. Similarly, with solution at $\lambda_k$ as the reference, we can keep screening for a number of $\lambda$-values: $\lambda_{k+1},\lambda_{k+2},...$ until it is beneficial to update the reference to the newer solution at some $\lambda_{K+B}$ and in this way all the discarded features are guaranteed to be zero. Adaptive EDPP is still a safe rule, but it is neither sequential nor one-step.

Similar to SEDPP, the Adaptive EDPP rule can be stated in the following theorem:

@@

For each $\lambda$-value in one update $\lambda_{k+1},...,\lambda_{k+B}$, the computation in \eqref{eq:aedpp1} or \eqref{eq:aedpp2} is the same, except for one quantity $\lambda_{k+b}$ ($\lambda_b$ in the case $k=0$). The most expensive computations are $x_j^Tr(\lambda_k)$, $x_j^Tx_*$ and $x_j^Ty$, for $j=1,2,...,p$. $x_j^Ty$ and $x_j^Tx_*$ are the same for all $\lambda$-values, $x_j^Tr(\lambda_k)$ are the same for all $\lambda$-values within the update and $x_j^T\hat{y}(\lambda_k)=x_j^Ty-x_j^Tr(\lambda_k)$, so once they are computed, they can be stored and reused to avoid more computation in that update. These two computations take $O(np)$ time and the rest computation in \eqref{eq:aedpp1} or \eqref{eq:aedpp2} take at most a lower order of $O(p)$ time.

Compared to one-step methods like BEDPP, after updating, the solution as the reference at $\lambda_k$ and the solution at some $\lambda_{k+b}$ in that update are usually relatively close, so screening at $\lambda_{k+b}$ using solution at $\lambda_k$ will remain powerful. For one-step method, as $\lambda_k$ gets further away from $\lambda_0$, the screening power can drop drastically and may not be any beneficial after that. Compared to sequential methods like SEDPP, the adaptive method only requires one $O(np)$ computation in one update, while SEDPP requires one $O(np)$ computation for each $\lambda$-value. The framework of adaptive safe screening can also be easily applied to other sequential safe rules, to form a family of adaptive safe rules to achieve the same improvement.

\subsection{Adaptive hybrid rule on top of adaptive safe rule}

>>>>

\subsection{Safe rules}
\label{sec:safe}

@@

\pb{Mention Sure Independence Screening \citep{Fan2008} and why we do not focus on it here; maybe in general, just introduce the idea of stochastic screening}

\section{Adaptive screening}
\label{sec:method}

% In this section, we will introduce the framework of adaptive hybrid rules, implement them with EDPP as the safe rule to form Ada-SSR-EDPP and propose a strategy to adaptively updating the reference.

\subsection{Adaptive safe rules}

The SEDPP rule \eqref{eq:sedpp} from Theorem 2.2 allows us to screen features at $\lambda_{k+1}$ using the solution at $\lambda_k$ as a reference. However, since the grid of $\lam$ values is arbitrary, the rule can be just as easily applied to screen solutions at $\lam_{k+2}, \lam_{k+3}, \ldots$.  The difference is that in typical SEDPP usage, we would update the reference solution at every step of path; here, we are proposing to re-use solutions for several values of $\lam$, only updating the reference when it is computationally beneficial to do so.  We refer to this modification as \emph{adaptive EDPP}.  Adaptive EDPP is still a safe rule, but it is neither sequential nor one-step.  A similar idea would be to use the reference solution in batches, each one being used for a fixed number of $\lam$ steps.  However, since an appropriate batch size is difficult to determine ahead of time, and also because the ideal size changes along the path, we find that an adaptive approach is better. \pb{Do we have evidence for this?}

The Adaptive EDPP rule can be stated in the following theorem, which simply restates the theorems from Section~\ref{sec:safe} in terms of this new paradigm in which a single reference solution at $\lam_k$ is used to calculate screening rules at $\lam_{k+2}, \ldots, \lam_{k+B}$.

@@

For each $\lambda$ value in $\lambda_{k+1},...,\lambda_{k+B}$, the computation in \eqref{eq:aedpp1} or \eqref{eq:aedpp2} is the same, except for the quantity $\lambda_{k+b}$ ($\lambda_b$ in the case $k=0$). The expensive computations in calculating these rules are $x_j^Tr(\lambda_k)$, $x_j^Tx_*$ and $x_j^Ty$, all of which must be calculated for $j=1,2,...,p$ and thus require $O(np)$ operations; all other computations require at most $O(p)$ operations. However, since $x_j^Ty$ and $x_j^Tx_*$ are the same for all $\lambda$ values and $x_j^Tr(\lambda_k)$ is the same for all $\lambda$ values within the block, once they have been calculated once, they can be stored and reused to avoid additional computation for all remaining steps in the block.  Note that $x_j^T\hat{y}(\lambda_k)=x_j^Ty-x_j^Tr(\lambda_k)$, so this term also can be calculated efficiently.

Compared to one-step methods like BEDPP, the reference solutions are continually updated.  As a result, the reference is always relatively close to the solution, and screening remains powerful over the entire path, unlike one-step methods where screening power drops drastically as $\lambda_k$ gets further away from $\lambda_0$. Compared to sequential methods like SEDPP, the adaptive method only requires $O(np)$ computations per update, while SEDPP requires $O(np)$ computations for each $\lambda$ value. The framework of adaptive safe screening can also be easily applied to other sequential safe rules, to form a family of adaptive safe rules.

\subsection{Adaptive hybrid rules}
